---
title: 'Statistical Inference Course Project: Simulation'
author: 'Kevin Tai'
output:
  pdf_document: default
---

# Overview

In this project we will investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Set lambda = 0.2 for all of the simulations. We will investigate the distribution of averages of 40 exponentials over a thousand simulations.

Specifically, our goals are to illustrate via simulation and associated explanatory text the properties of the distribution of the mean of 40 exponentials. We will

1. Show the sample mean and compare it to the theoretical mean of the distribution.
2. Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.
3. Show that the distribution is approximately normal.

# Simulations
First, we perform the simulations. We can do this in a manner similar to the one presented in the lectures. Ultimately, we will produce the variable means which contains 1000 averages of 40 draws from an exponential distribution.

```{r}
library('ggplot2')
lambda <- 0.2
n <- 40
trials <- 1000
m <- matrix(rexp(n * trials, lambda), trials, n)
means <- rowMeans(m)
```

To get a quick sense of the data we've simulated, let's take a look at a histogram of the sample means.

```{r, message=FALSE, warning=FALSE}
library('ggplot2')
qplot(means, geom='histogram', main='Sample Mean of Exponential Distribution',
      ylab='Frequency', xlab='Means')
```

# Sample Mean vs. Theoretical Mean
First let's compare the sample means with the theoretical mean.

Based on the lectures, we know that the sample mean should be equal to the population mean. With lambda set to 0.2, the theoretical mean should be 5. Let's look at the cumulative means to see if the mean of the sample means converges to the theoretical. The red line indicates the theoretical mean and the black line is the cumulative means. 
```{r}
mean(means)
```

We can see over a large number of trials, the sample means converge to an number which is pretty close to the theoretical. 

```{r, warning=FALSE}
cum_means <- data.frame(x=1:trials, y=cumsum(means) / (1:trials))
# Create the cumulative means plot
ggplot(cum_means, aes(x=x, y=y)) + geom_line() + ylim(c(4.75, 5.8)) +
  ggtitle('Cumulative Means of the Means') + ylab('Cumulative Means') + xlab('Index') + 
  geom_abline(slope=0, intercept=5, color='red')
```

# Sample Variance vs. Theoretical Variance
Let's also compare the sample variance with the theoretical variance. We can take the original matrix and compute the variance for each 40-draw sample. The theoretical standard deviation is 1/lambda, so the theoretical variance should be 25.

```{r}
variances <- apply(m, 1, var)
mean(variances)
```

The average of the sample variances is pretty close to the theoretical. Like before, we can compute the cumulative average of the variances.

```{r}
cum_vars <- data.frame(x=1:trials, y=cumsum(variances) / (1:trials))
# Create the cumulative means plot
ggplot(cum_vars, aes(x=x, y=y)) + geom_line() +
  ggtitle('Cumulative Means of the Variances') + ylab('Cumulative Means') + xlab('Index') + 
  geom_abline(slope=0, intercept=25, color='red')
```

# Distribution
According to the Central Limit Theorem, we know that this distribution ought to be close to a normal distribution. We can adjust the sample means to adjust it to a standard normal and compare.

```{r}
adj_means <- (means - 5) / (5 / sqrt(n))
ggplot(data.frame(value=adj_means), aes(x=value)) +
  geom_histogram(alpha = .20, binwidth=.3, colour = "black", aes(y = ..density..)) + 
  stat_function(fun=dnorm) + ggtitle('Sample Means Distribution')
```

We can see that following normalization, the plot is fit by the standard normal pretty well.